### 2024-12-03 02-methods.Rmd

```{r setup-methods, include = FALSE}
require(kableExtra)
```

# Methods {#sec:methods}

## AI Tutor

To assess the impact of LLMs on students' critical thinking, we developed an _AI Tutor_, a web-based application built using Python with the Flask framework. The application integrates OpenAI APIs, allowing students to interact with the GPT-4.0 model while performing different educational tasks. This app's key features include:

- **User authentication**: A secure and anonymous login system for students.
- **Web Survey**: A web interface to survey students and where they could perform simple tasks online, such as writing a short essay or performing basic calculations.
- **AI Chatbot**: A chatbot powered by GPT-4 provided students with personalised assistance and tutoring during the session. A new chatbot instance was associated with each question for each student, isolating each conversation from interactions in previous questions. The chatbot was available to students only for specific questions or tasks under the researchers' control.
- **Data logging**: A SQL database storing students' conversation logs with the AI chatbot, including texts, timestamps, and survey responses. 
- **Randomised assignment**: A system to randomly assign registered students to different versions of the AI chatbot or treatment groups to explore the causal impact of various configurations on students' interactions and performance.

<!-- ![AI Tutor Screen Shot](pictures/placeholder.jpg) -->

## Participants

We recruited students  between 14 and 16 years old enrolled in secondary education from two schools: one in Brussels (Belgium) and one in Seville (Spain). The experiment was conducted during school hours at the schools (Fig. \@ref(fig:classrooms)). Both schools are bilingual institutions, with English as a primary language, ensuring participants had a high English proficiency. The AI Tutor was multilingual, allowing students to interact with either language. The experimental instructions were in English. However, if needed, students could translate the instructions using the browser's automatic translation service.

Recruiting students was conducted after receiving ethical approval from the European Commission's internal Ethical Review Board, ensuring that the consent procedures, data protection requirements, and the experimental protocol complied with local laws and were safe for the participants. The data privacy protection protocol was approved by the Commission's data protection officer. As an additional safeguard, given the minor age of the participants, we required their parents or legal guardians to provide us with written consent, allowing their children to participate in the study. The students also had to give their assent to participate by completing an online consent form.

![Classrooms used for the experiment in Seville (**A**) and in Brussels (**B**).](assets/pictures/classrooms.pdf){#fig:classrooms}

## Experimental sessions

About ten experimental sessions were conducted at the schools between November 11 and 12, 2023, in Brussels and February 8 and 9, 2024, in Seville. Each session took about two hours. In the first 45-60 minutes, students received a personal computer and were asked to perform multiple tasks online using the AI tutor, including answering a questionnaire. In the following 45-60, there was a group discussion on how the students found the interactions with the AI tutor and a more general debate on how they perceived the potential benefits and drawbacks of integrating LLMs in the classroom. The results of the group discussions are not discussed in this paper; however, they were used as material for interpreting the results of the experimental study.

## Experimental Conditions

Figure \@ref(fig:design) illustrates the two randomised manipulations in our experimental design: (1) AI step-by-step reasoning and (2) Socratic vs non-Socratic AI.

![The experimental design involves two manipulations: **A.** Socratic vs Non-Socratic; **B.** AI solution with explanation vs AI solution (without explanation)](assets/pictures/experimental_design.pdf){#fig:design}

#### AI step-by-step reasoning {-}

The first manipulation focuses on a task in which students are asked to estimate the value of coins in a jar (see Section \@ref(sec:si-ai-explanation) for the details).[^jar] Specifically, we used the coin jar from Steiner's experiment [@steiner2015turns], aimed originally at assessing Internet users' guessing accuracy. For this intervention, we varied the AI’s response by providing either a complete answer, which included both an estimated value and a step-by-step explanation generated by the AI tutor, or a partial answer that provided only the estimate without additional details. In both conditions, all participants received the same estimated value ($213), but those in the full-answer condition also viewed an explanation outlining the step-by-step reasoning the AI used to arrive at this estimation based on the jar image.[^jar-gpt4]

[^jar]: This is a common experimental activity in economics, especially in the context of auction theory [@thaler1988anomalies]. The task is ideal in our setting because it requires participants to guess based on limited information, thus creating a situation where AI-generated assistance could potentially influence their decisions.

[^jar-gpt4]: Using GPT 4.0, we uploaded the image of the coin jar asking for an estimate of the value of coins for ten times. We then selected the median response for the experiment.

This setup allows us to examine how AI-provided explanations affect students' performance and their perceptions of the AI's accuracy. Specifically, we focus on three outcome variables: (1) the propensity to update their initial guess and size of their updates, (2) the accuracy of students' final estimations, measured as the (absolute) difference between their guesses and the actual coin value, and (3) students' perceived accuracy of the AI, rated on a scale from low to high. We also asked participants (4) to rate the perceived accuracy of the mean guess among 600 people ($596), which exaggerated the correct value, as reported in the original article [@steiner2015turns].

#### Socratic vs Non-Socratic AI {-}

The second manipulation involved randomly assigning students to one of two different types of AI tutors: Socratic or non-Socratic AI tutors. Both tutors were powered by the same underlying large language model (GPT-4). Still, each was instructed with different "system messages" to create different behaviours, as illustrated in Table \@ref(tab:prompts).[^prompts] For the Socratic tutor, the system message asked the model to engage students with open-ended, thought-provoking questions, encouraging them to think critically about their responses. In contrast, the non-Socratic AI tutor was instructed to provide concise, direct answers without necessarily engaging in deeper dialogue or posing further questions.

[^prompts]: An AI's system message guides how the AI interprets the conversations by setting paramters for interaction.

```{r prompts}
df_prompts <- data.frame(
  "AI Tutor" = c("Socratic", "Non-Socratic"),
  "Prompt instruction" = c("You are a Socratic tutor. You always answer using the Socratic style, asking just the right questions to help students learn to think for themselves, breaking down the problem into simpler parts until it's at the right level for them. You provide concise information and explanations understandable for 8th to 10th grade students.", "You are a didactic tutor. You provide concise information and explanations understandable for 8th to 10th grade students.")
)
kbl(df_prompts, booktabs = TRUE, caption = "Prompt instructions associated with the AI Tutor's treatments") |>
  kableExtra::column_spec(2, "4in")
```

This setup allowed us to investigate the impact of different pedagogical AI tutoring approaches on students' performance and perceptions. Specifically, we asked students to use the AI tutor while performing three tasks: guess an unknown quantity ("How much water in litres do students consume at our school each week?"); express an opinion and write a short essay ("What is your opinion about the effect of social media on teenagers?"); respond to physics questions on how sound propagates in different media ("In which of the following materials does sound travel faster?"). These questions enabled us to assess the AI tutor's impact on students' learning and problem-solving abilities.

We focused on two primary metrics: (1) confidence in their answers, (2) perceived usefulness of interacting with the AI tutor, and (3) correctness of their answers (for task 3). Specifically,  students were asked, "How confident are you that the answer you provided is accurate?" on a five-point scale ranging from "not confident at all" to "very confident." They were also asked "How helpful was it to interact with the AI tutor?" This was also rated on a five-point scale, from "Not at all helpful" to "Very helpful." Finally, we also assessed the correctness of the response, whenever a correct answer was available (i.e., sound propogation questions).

## Background Measures

As an initial step in our analysis, and to better understand students' awareness of and attitudes toward AI, we developed a brief four-question "AI in Education Attitudes Scale." This scale was adapted from the broader AI Attitudes Scale proposed by Schepman and Rodway [@schepman2020initial]. These questions included: Do you agree or disagree that society will benefit from a future of AI? Do you agree or disagree that AI is dangerous? Do you agree or disagree that AI will foster students’ learning in the future? Do you agree or disagree that AI is often misused by students?

Finally, to consider participants' academic performance and skills, participants were also asked about their school grades and how often they complete their homework assignments on time and what factors affect their ability to do so. Self-reported experience using ChatGPT (the main LLM available at the time of the study) was also asked, including a question on their estimation of how many of their peers use ChatGPT for their homework.

Section \@ref(sec:questionnaire) presents the complete questionnaire.