### 2024-12-03 80-references.Rmd

\newpage

# Acknowledgements {-}

The contents of this publication do not necessarily reflect the position or opinion of the European Commission or the Joint Research Centre (JRC). We thank the colleagues in the JRC's Digital Economy unit for their support and generous collaboration. Special thanks for the Data Privacy Coordinator of the JRC, who helped us with the data privacy statement, ensuring the highest safeguard to the students participating in the study. Many thanks go to the schools, including their leadership teams, parents and pupils who participated in the study. All errors are our own.

\newpage

# References

::: {#refs}
:::