### 2024-12-03 99-questionnaire.Rmd

\newpage

## Questionnaire {#sec:questionnaire}

1. Do you agree or disagree that society will benefit from a future of AI?
   - Strongly Agree
   - Agree
   - Neutral
   - Disagree
   - Strongly Disagree

2. Do you agree or disagree that AI is dangerous?
   - Strongly Agree
   - Agree
   - Neutral
   - Disagree
   - Strongly Disagree

3. Do you agree or disagree that AI will foster students’ learning in the future?
   - Strongly Agree
   - Agree
   - Neutral
   - Disagree
   - Strongly Disagree

4. Do you agree or disagree that AI is often misused by students?
   - Strongly Agree
   - Agree
   - Neutral
   - Disagree
   - Strongly Disagree

5. How easy or difficult would it be for you to explain how sound waves transfer in the air or other materials?
   - I could do this easily
   - I could do this with a bit of effort
   - I would struggle to do this on my own
   - I couldn't do this on my own

6. How easy or difficult would it be for you to write about the impact of technology on teenagers’ well-being?
   - I could do this easily
   - I could do this with a bit of effort
   - I would struggle to do this on my own
   - I couldn't do this on my own

7. How easy or difficult would it be for you to guess how many litres of water are in an Olympic swimming pool?
   - I could do this easily
   - I could do this with a bit of effort
   - I would struggle to do this on my own
   - I couldn't do this on my own

8. You see a jar filled with different coins; you must guess the total value of coins in US dollars.

9.  We asked the same question about the jar's coins value to an <b>AI system</b> that can understand and process visual information. The AI guessed $213. How accurate is this guess? [IF in AI EXPLANATION TREATMENT, ADD EXPLANATION HERE]
    - Correct
    - Mostly Correct
    - Partially Correct
    - Mostly incorrect
    - Incorrect

10. We asked the same question to a group of 600 people of various backgrounds. People's mean guess was $596. How accurate is this guess?
    - Correct
    - Mostly Correct
    - Partially Correct
    - Mostly incorrect
    - Incorrect

11. Given the information from the AI ($213) and the people's average ($596), what is your final guess of the value of coins in the jar?

12. [Task 1] How much water in litres do students consume at our school each week? Interact with the AI tutor at the bottom of this page before answering. [Randomly assign SOCRATIC / NON-SOCRATIC AI Tutor]

13. How confident are you that the answer you provided is accurate?
    - Very confident
    - Confident
    - Neutral
    - Not very confident
    - Not confident at all

14. How helpful was interacting with the AI tutor?
    - Very helpful
    - Helpful
    - Neutral
    - Not very helpful
    - Not helpful at all

15. [Task 2.] Experts say that social media can have either a positive or a negative impact on students' well-being. What is your opinion about the effect of social media on teenagers? There is no correct or wrong answer.
    - Very Positive
    - Positive
    - Neutral
    - Negative
    - Very Negative

16. Now, interact with the AI tutor at the bottom of this page before answering. [SOCRATIC / NON-SOCRATIC]. [Repeat question 15.]
    - Very Positive
    - Positive
    - Neutral
    - Negative
    - Very Negative

17. Write a well-reasoned 600-character essay critically examining this topic. Write an introductory paragraph with background information, argumentation with as many convincing arguments or facts as possible, and a brief conclusion.

18. How confident are you that the arguments or facts in your essay are accurate?
    - Very confident
    - Confident
    - Neutral
    - Not very confident
    - Not confident at all

19. How helpful was interacting with the AI tutor before writing the essay?
    - Very helpful
    - Helpful
    - Neutral
    - Not very helpful
    - Not helpful at all

20. Can you explain why this interaction was helpful or wasn't? [TEXT]

21. [Task 3] Does sound travel faster in water than air? And if so, why?
    1.  Sound travels slower in water due to its higher density
    2.  Sound travels faster in water due to its higher density [correct answer]
    3.  Sound travels at the same speed in both water and air, regardless of density
    4.  Sound travels slower in water due to its lower density
    5.  Sound travels faster in water due to its lower density

22. Now, interact with the AI tutor at the bottom of this page before answering. [SOCRATIC / NON-SOCRATIC] [Repeat question 21.]

23. How confident are you that the answer you provided is accurate?
    - Very confident
    - Confident
    - Neutral
    - Not very confident
    - Not confident at all

24. How helpful was interacting with the AI tutor?
    - Very helpful
    - Helpful
    - Neutral
    - Not very helpful
    - Not helpful at all

25. [Verification question] In which of the following materials does sound travel faster?
    - Rubber
    - Cold Air
    - Warm Air
    - Gold
    - Water [correct answer]

26. What is your average grade at school?

27. How many hours per day do you spend completing homework assignments?

28. How often do you complete your homework assignment on time?

29. What single factor contributes the most to your ability to complete homework assignments on time?

30. Have you ever used ChatGPT before today?

31. In your opinion, how many of your classmates are using ChatGPT for homework?