### 2024-12-16 00-abstract.Rmd
submission

# Highlights {-}

- An experiment was conducted where K-12 students used AI chatbots to solve school-related problems.

- Two AI chatbot approaches were tested: one provided incremental guidance to encourage critical thinking ("Socratic"), while the other offered immediate solutions ("non-Socratic").

- Results showed that AI-generated explanations improved students' performance over solutions without them, highlighting the value of AI-generated guidance.

- Participants engaged more frequently with the Socratic AI, though this did not result in improved performance compared to the non-Socratic AI.

- While students found interactions with the AI useful, they perceived the Socratic AI as less helpful overall.

- The findings highlight challenges in designing AI tutors that effectively foster critical thinking while maintaining student satisfaction, raising concerns about their adoption in educational settings.

# Abstract {-}

How does integrating large language models (LLMs) into classroom activities influence students' learning? To address this question, we conducted a randomized experiment comparing two distinct chatbot approaches: one designed to encourage critical thinking through incremental guidance ("Socratic AI") and another providing immediate solutions ("non-Socratic AI"). The study involved students aged 14 to 16, who engaged in various school-related tasks under these experimental conditions. Students' attitudes were measured through self-reported surveys. Results indicated that AI-generated explanations significantly improved students' performance over solutions provided without such explanations, highlighting the benefits of step-by-step guidance. The Socratic AI approach fostered significantly greater engagement and interaction. However, it did not achieve significant improvements in learning, and a higher fraction of students perceived it as less helpful. Furthermore, despite students generally perceiving AI assistance as beneficial, they exhibited limited retention, failing to apply learned concepts to new situations when we removed AI assistance. These findings contribute to the ongoing debate on integrating LLM-powered chatbots in education, highlighting key challenges in designing AI tutors that effectively foster critical thinking while maintaining student satisfaction, raising concerns about their adoption in educational settings.

<!-- How does integrating Large Language Models (LLMs) into classroom activities influence students' learning? We conduct a randomised experiment to examine whether LLMs' interactions and step-by-step reasoning improve students' performance. Specifically, we compare chatbots designed to encourage critical thinking through incremental guidance ("Socratic AI") against chatbots providing immediate solutions ("non-Socratic AI"), evaluating the effects on students' engagement, interactions, and learning outcomes. Students aged 14 to 16 participated in problem-solving and writing tasks under different experimental conditions. We evaluated their performance through conversation logs and test scores on various school-related tasks. Attitudes were measured using self-reported surveys.  -->

**Keywords**: artificial intelligence; large language models; learning; education policy; experiment, k-12

<!-- Competing Interests: Authors are required to disclose financial or non-financial interests that are directly or indirectly related to the work submitted for publication. Please refer to “Competing Interests and Funding” below for more information on how to complete this section. -->

<!-- 
Cite references in the text by name and year in parentheses. Some examples:

Negotiation research spans many disciplines (Thompson, 1990).
This result was later contradicted by Becker and Seligman (1996).
This effect has been widely studied (Abbott, 1991; Barakat et al., 1995; Kelso & Smith, 1998; Medvec et al., 1999). 
-->