### 2024-12-16 main_report.Rmd

---
title: |
  Exploring the Role of AI Chatbots in K-12 Education: A Comparative Study of Socratic and Non-Socratic Approaches

author:
  - Andrea Blasco:
      institute:
        - jrc
      correspondence: "yes"
      email: andrea.blasco@ec.europa.eu
  - Vicky Charisi:
      institute: [jrc, harvard]
      correspondence: "yes"
      email: vcharisi@law.harvard.edu

institute:
  - jrc: European Commission, Joint Research Centre, Brussels, Belgium
  - harvard: Harvard University, Berkman Klein Center, Cambridge, MA, USA

date: |
  `r format(Sys.Date(), 'This version: %B %d, %Y')`

toc: false
fontsize: 11pt # 11
linestretch: 2 # 1.25
colorlinks: true
link-citations: true
graphics: true
tables: true
mainfont: Times New Roman
geometry: margin=2cm
bibliography: bibliography/refs.bib
always_allow_html: true

header-includes: |
  ```{=latex}
  % Break sections into new page
  \usepackage{titlesec}
  \newcommand{\sectionbreak}{\clearpage}

  % Table caption 
  \usepackage{caption} 
  \captionsetup[table]{skip=8pt}

  % Table packages
  \usepackage{longtable}
  \usepackage{array}
  \usepackage{multirow}
  \usepackage{wrapfig}
  \usepackage{float}
  \usepackage{colortbl}
  \usepackage{pdflscape}
  \usepackage{tabu}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage[normalem]{ulem}
  \usepackage{makecell}
  \usepackage{xcolor}

  % Move all figures at the end
  \usepackage[nolists, nomarkers]{endfloat}
  \DeclareDelayedFloatFlavour*{longtable}{table}

  %\graphicspath{}
  ```
---

<!-- # The Impact of Large Language Models on Students: A Randomised Study of Socratic vs. Non-Socratic AI and the Role of Step-by-Step Reasoning -->

```{r setup, include = FALSE}
library(patchwork)
library(figpatch)
knitr::opts_chunk$set(echo = FALSE, dev = "pdf")
```

```{r child=list.files(pattern = "[0-9]+.*.Rmd", full = TRUE)}
```