### 2025-05-23 Reviews/round_1/rebuttal_point_by_point.md
edits

# Rebuttal to Reviewer Comments – Round 1

Thank you for the constructive feedback. Below is our detailed, point-by-point response to the reviewers' comments. Reviewer comments are in **bold**, and our responses follow in regular text with manuscript changes noted where applicable.

---

## Reviewer 1

> This paper investigates whether and how the integration of large language models (LLMs) into classroom activities influence students’ learning. Although the topic, questions, and results are crucial, the manuscript requires revision before publication. Therefore, I recommend a major revision, and I am willing to review the revised manuscript. I have provided a list of comments below to assist the authors in transforming their paper into a manuscript worthy of publication.

### Comments

> The abstract sound great.

Thanks!

> The literature review provides a broad context for AI’s potential; however, it lacks a strong connection to recent studies that address AI’s evolving role in educational contexts, critical thinking, and argumentative writing. The paper could enhance its theoretical foundation by engaging more deeply with contemporary research on AI’s impact on students' critical thinking, argumentation and reasoning. There are much more recent studies by scholars such as Kazem Banihashem and that you can consult with.

We did ...

> Please justify why the participants come from Brussels (Belgium) and one in Seville (Spain)? Why these two countries and these two cities? What is the argument here?

We recruited two schools in different countries to increase the external validity of our study. The country choice was by convenience as the authors lived in Brussels and Seville at the time of the experiment. We sought schools with similar curricula and contacted the schools through personal connections. The selected schools were both international with students from comparable socio-economic backgrounds, which was ideal for conducting the experiment.

> More information on the method section, the context, the design, and the participant characteristics are needed to give the readers a clear description of the study. Please elaborate on the background information of the characteristics of the participants that could influence the way learners engage with AI and critical thinking and argumentation. Acknowledging these limitations can guide future research in this area.

We have included more information ....

> The results show promising outcomes. However, the findings of this study are derived from self-reported data collected from participants through surveys. However, it is important to note that the literature suggests caution in heavily relying on such self-reported data, as perception does not always equate to actions (see: https://urldefense.com/v3/__https://doi.org/10.1080/02602938.2024.2345669__;!!DOxrgLBm!DU2xLNvI0Ke45t43rGQAI3BYa8bni0Io2JesBwINodtn8kzLHvspBE2my62mGZlAoKzFv6RFxgUv6FWyBQIbCg9ObQu4SA$ ). Therefore, this must be acknowledged as a limitation of the study, along with suggestions for future research to address this issue.

We have discussed this limitation in the conclusions ...

> The depth and width of the discussion section could be improved by incorporating relevant and recent literature. This would provide more theoretical insights and add value to the study by linking the findings to the broader literature.

Many thanks we have ...

## Reviewer 2

> The article deals with a current and pressing topic. Since AI-based Large Language Models have become freely accessible, the way in which students learn has changed rapidly. There are new opportunitiesbut also risks for didactic methods. The article addresses the use of an AI tutor to support students' learning and understanding. An experimental setup is used to investigate the conditions under which an AI tutor enhances learning, how an AI tutor influences students' confidence in their own performance, and whether an AI tutor is perceived as helpful by students.

### Comments

> The article presents three research questions. The first question should be expressed with greater precision. Of particular note, the second part of the research question is not addressed by the selected experiments. The second research question, which comprises two distinct inquiries, should be subdivided to enhance its clarity and comprehensibility. The research questions are derived from existing literature. However, this section is rather brief. Notably, no hypotheses have been formulated or derived from the literature. Only one hypothesis has been stated explicitly. However, this has been done in the results section only, contrary to scientific conventions.

Clarified the research questions.

> In the Methods section and in the Supplementary Materials, the experimental procedure is largely described in a comprehensive and reproducible manner. However, the methods/experiments/questionnaire items are not related to the previously stated research questions or hypotheses. This makes it unnecessarily difficult for the reader to assess the validity of the approach chosen to address the research questions. To enhance comprehensibility, it would be beneficial to explicitly label the independent and dependent variables, as well as the control variables collected for each research question/hypothesis.

Ok.

> The results section first describes the group of subjects. For one experimental intervention (socratic AI tutor), it is reported whether the experimental group differed significantly from the control group. However, this information is missing for the other experimental intervention (AI explanation). In the remainder of the section, the analyses of the dependent variables are presented in a largely scientific manner. Notably, for the first experimental manipulation, a dependent variable has been listed in the methods section (propensity to update initial guess). However, no results are reported for this variable, and I do not understand what was done with this data or why it wasn't used. In the second experimental manipulation (socratic tutor), variables are reported that have not been described in the methods section. These new variables that measure the interaction with the AI tutor are only (and very briefly) motivated here. They should be motivated in the Introduction and included in the Methods section.

Ok.

> Contrary to usual conventions, some results with a significance level greater than 0.1 are reported as significant. In addition, in some cases, the language used is not specific/detailed enough to correctly understand the analyses (Which analysis uses exactly which data or answers from the questionnaire?). In some cases, the methods used are only briefly named and may therefore not be reproducible (in particular bootstrap method, in some cases also regression models - which variables exactly have been included?).

> The discussion addresses the research findings and identifies some methodological limitations. Since there was little specific literature on the specific research questions in the Introduction, the discussion of the results in the context of the existing literature is also very brief. However, it is particularly striking that the entire discussion does not contain a single reference/citation. In my opinion, this would be essential for a scientific publication.

Ok.

> In my opinion, some interpretations of the results go too far. In particular, the data on learning and retention effects can only give first indications, because each type of task was completed exactly once by the students. The database is therefore too small to provide generalizable results. Nevertheless, the article contains interesting results, especially on the interaction/acceptance of a socratic AI tutor by people who had prior experience with chatGPT.

ok.

> The methodological discussion contains important points but could be more detailed. In particular, the validity of the methods used should be discussed (was the question about transferability of knowledge appropriate (sound/density)? ). Further, it is not discussed whether the setting at different schools in two different countries could have influenced the results.

OK

> Overall, the article contains good and important research approaches and first interesting results, but the scientific reporting should be significantly improved.
(If you wish, you can find more detailed comments in the annotated PDF.)

ok.