### 2025-05-24 03-results.Rmd

```{r setup-results, include = FALSE}
require(dplyr)
require(kableExtra)
require(tidyr)
library(ordinal) #clm()
library(ggplot2)
library(ggdist)
library(ggrepel)
library(ggtext)
library(knitr)
library(patchwork)
set.seed(4881)

opts_chunk$set(echo = FALSE, message = FALSE)

source("analysis/R/text_functions.R")
source("analysis/config.R")

tutorgpt <- readRDS("analysis/data/processed/tutorgpt.rds")
chats <- readRDS("analysis/data/processed/chats.rds")
tutorgpt_wide <- readRDS("analysis/data/processed/tutorgpt_wide.rds")

geom_coef <- function() {
  list(
    scale_color_manual(values = c("navy", "orange")),
    geom_vline(xintercept = 0, linetype = "dashed"),
    geom_linerange(linewidth = 3, alpha = 0.2), 
    geom_point(size = 3),
    guides(colour = guide_legend(position = "inside"))
  )
}
```

# Results {#sec:results}

## Overview of Student Demographics

A total of 122 students participated in the study, 64 in Brussels and 58 in Seville. The sample was gender balanced. Over half of the students reported B+ grades and prior use of ChatGPT, with boys more likely to report ChatGPT experience. Around 50% reported to always complete homework on time, with a minority reporting less than always, with factors like time management (17% in Brussels, 25% in Seville) and material comprehension (56% in Brussels, 28% in Seville) affecting completion. Students' self-efficacy varied by task, with 30% of the students feeling they could "easily" write a technology essay (task 2) but only 6% and 11% feeling confident about explaining sound propagation (task 3) or solving a numerical estimation task such as guessing the litres of water in a pool (task 1). Thus, the sample was substantially homogeneous in terms of demographics but diverse in terms of self-efficacy.

Table \@ref(tab:summary-statistics) illustrates the distribution of student characteristics across treatment groups alongside the results of separate Fisher's Exact Tests on the association with treatment assignment. The sample characteristics were generally balanced across treatment groups. Only one variable out of ten -- i.e., self-reported difficulties in homework -- was statistically associated (p = 0.05) with the Socratic/Non-Socratic assignment, while no significant associations were found for the AI-reasoning assignment. Thus, only one out of twenty Fisher's exact tests showed a significant result, indicating a good balance across treatments.

```{r summary-statistics, echo = FALSE, warning=FALSE, message=FALSE}
tutorgpt_rename <- tutorgpt %>%
  dplyr::select(
    treatment_tutor_id, 
    treatment_explain_id,
    "Gender" = gender,
    "Has used ChatGPT" = has_used_before,
    "Location" = location,
    "Grades" = top_grades,
    "Homework weekly study" = hours_per_week,
    "Homework on time" = on_time_completion,
    "Homework difficulties" = factors_contributing,
    "Self-efficacy (sound)" = q5_self_physics,
    "Self-efficacy (essay)" = q6_self_technology,
    "Self-efficacy (guess)" = q7_self_swimming_pool,
  )

table_socratic <- tutorgpt_rename %>%
  select(-treatment_explain_id) %>% 
  tidyr::pivot_longer(-treatment_tutor_id) %>%
  mutate(pval = xtabs(~ value + treatment_tutor_id) %>% 
                fisher.test() %>% 
                .$p.value,
        .by = name)  %>% 
  dplyr::count(name, value, pval, treatment_tutor_id) %>%
  na.omit() %>%
  dplyr::mutate(percent = 100 * n / sum(n), 
                .by = c(name, treatment_tutor_id)) %>%
  tidyr::pivot_wider(names_from = treatment_tutor_id,
                     values_from = c(percent, n),
                     values_fill = 0) %>%
  dplyr::mutate(totals = n_non_socratic + n_socratic)

table_explain <- tutorgpt_rename %>%
  select(-treatment_tutor_id) %>% 
  tidyr::pivot_longer(-treatment_explain_id) %>%
  dplyr::mutate(pval_explain = xtabs(~ value + treatment_explain_id) %>% 
                               fisher.test() %>% 
                                .$p.value, 
                .by = name)  %>% 
  dplyr::count(name, value, treatment_explain_id, pval_explain) %>%
  na.omit() %>%
  dplyr::mutate(percent = 100 * n / sum(n), 
                .by = c(name, treatment_explain_id)) %>%
  tidyr::pivot_wider(names_from = treatment_explain_id,
                     values_from = c(percent, n),
                     values_fill = 0)

table_socratic %>% 
  left_join(table_explain, by = c("name", "value")) %>% 
  mutate(name = ifelse(row_number() == 1, name, ""), 
         pval = ifelse(row_number() == 1, sprintf("%0.2f", pval), ""),
         pval_explain = ifelse(row_number() == 1, sprintf("%0.2f", pval_explain), ""),
         .by = name) %>%
  dplyr::select(
      name, value,
      percent_socratic, percent_non_socratic, pval,
      percent_explain_yes, percent_explain_no, pval_explain,
      totals,
  ) %>% 
  kableExtra::kbl(
    col.names = c(
      "Name", "Value",
      "Yes", "No", "p.",
      "Yes", "No", "p.",
      "N"
    ),
    digits = 0,
    booktabs = TRUE,
    caption = "Sample Characteristics",
    linesep = if (knitr::is_latex_output()) "\\hline" else "",
  ) %>% 
  kableExtra::kable_styling() %>% 
  kableExtra::add_header_above(c(" " = 2, "Socratic (%)" = 3,
                                 "AI Explain (%)" = 3, " "))
```

## Attitudes Towards AI

As illustrated in Figure \@ref(fig:attitudes-AI), students expressed conflicting views on the role of AI in education. On the one hand, a majority felt that AI is often misused by students (65%) and potentially dangerous (57%). On the other hand, most students also anticipated that AI would enhance student learning in the future (59%) and contribute positively to society (65%). These findings suggest that most students in our sample are aware of the risks of misuse and safety with a prevailing optimism that AI could support educational growth and societal progress.

```{r attitudes-AI, echo = FALSE, fig.cap = "Student attitudes on AI in education"}
agree_labels <- c(
  strongly_agree = "S-Agree",
  agree = "Agree",
  neutral = "Neutral",
  disagree = "Disagree",
  strongly_disagree = "S-Disagree"
)

palette <- c(
  "#ca0020", "#f4a582", 
  "#d1e5f0", "#92c5de", "#0571b0"
)

# Calculate response percentages
attitudes_df <- tutorgpt %>%
  dplyr::select(
    "Society will benefit from a future of AI?" = q1_ai_future,
    "AI is dangerous" = q2_ai_danger,
    "AI is often misused by students" = q4_ai_misuse,
    "AI will foster students' learning in the future" = q3_ai_learning
  ) %>% 
  tidyr::pivot_longer(cols = everything()) %>%
  dplyr::count(name, value) %>% 
  dplyr::mutate(percent = 100 * (n + 1) / (sum(n) + 5), .by = name)

# Create plot 
attitudes_df %>% 
  ggplot() + 
  aes(
    x = percent,
    y = factor(value, levels = names(agree_labels)),
    label = paste0(round(percent), "%"),
    fill = factor(value, levels = names(agree_labels)),
  ) +
  geom_col(width = 0.8, show.legend = FALSE) +
  geom_text(aes(hjust = ifelse(percent < 5, -0.5, 0.5)),
            position = position_stack(vjust = 0.5)) +
  facet_wrap(~ wrap_text(name), scales = "free") +
  scale_y_discrete(labels = agree_labels) +
  scale_fill_manual(values = rev(palette)) +
  theme(
    axis.line.x = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_blank(),
    panel.spacing = unit(2, "lines")
  ) + 
  labs(
    x = NULL, y = NULL, 
    caption = "Source: Full sample (n = 122)",
    title = "Student Attitudes on AI in Education",
    subtitle = "% of students who agree or disagree with the statement:",
  )
```

## The Impact of AI Step-by-step Reasoning Exposure

We tested whether showing students AI-generated step-by-step reasoning affected how they used the AI's prediction (RQ1), compared to just seeing the prediction without any reasoning. For each student $i = 1, \cdots, 122$, we computed the absolute error as the absolute difference between student $i$'s guess, $\text{Guess}_i$, and the actual value of coins:
$$
  \text{Absolute Error}_i = \mid \text{Guess}_i - \text{Actual Value Coins} \mid,
$$
where a smaller absolute error indicated greater accuracy.

<!-- The absolute error distribution in our sample was positively skewed, complicating the testing for mean differences. Consequently, we shifted our analysis towards treatment differences in the median absolute error, as illustrated in Figure \@ref(fig:boot-accuracy).  -->

Figure \@ref(fig:boot-accuracy) illustrates that students' prediction accuracy was positively skewed in both treatment groups, with a greater accuracy for students exposed to AI reasoning. To estimate confidence intervals for the median difference in accuracy betweeen the groups, we used a nonparametric bootstrap approach.[^outlier] This procedure involved resampling participants' absolute errors (n = 1,999) to build a distribution of medians for each group. The 5th and 95th percentiles of the resulting distribution were used to construct a 90% confidence interval for the difference in accuracy between groups. The interval ranged from 0 to 70, indicating a greater accuracy (lower error) for the students exposed to AI-generated reasoning (one-sided, p < 0.05).

[^outlier]: We detected one outlier in the data, which was removed from the bootstrap analysis: one student provided an unusual and notably high guess of 6969. However, this value deviated substantially from both the overall distribution and the student's initial guess, suggesting a mistake rather than an estimate.

```{r boot-accuracy, echo = FALSE, fig.width = 7, fig.height = 3, fig.cap = cap}
cap <- "Comparison of Students' Absolute Error in Estimating Coin Jar Value with and without AI Explanations. All students received the AI-generated estimate of $213 (the correct value was $379.54). Still, those in the AI reasoning treatment also viewed the AI-generated step-by-step explanation for the estimation. Exposure to the AI-generated explanation significantly reduced the students' median absolute error." # nolint

coins_value <- 379.54

# Regression
n_boot <- 1999
palette <- c("#d95f02", "#7570b3")

# # For each bootstrap sample we compute the sample median, boot_med
resample_median <- function(n_boot, x) {
  mat <- replicate(n_boot, sample(x, replace = TRUE))
  apply(mat, 2, median)
}

bootstrap_accuracy <- tutorgpt %>%
  filter(q12_guess_coins_final < 6e3) %>% 
  mutate(abs_err = abs(q12_guess_coins_final - coins_value)) %>%
  reframe(
    .by = treatment_explain_id,
    boot_med = resample_median(n_boot, abs_err),
    sim_id = seq(n_boot)
  )

bootstrap_accuracy_diff <- bootstrap_accuracy %>% 
  summarise(diff = diff(boot_med), .by = sim_id)

# Test diffrence
alpha <- 0.1
diff_cis <- bootstrap_accuracy_diff %>%
  reframe(quantile(diff, p = c(alpha / 2, 1 - alpha / 2)))

# # Compute p-values two-sided test
p_value <- bootstrap_accuracy_diff %>% 
  with(1 - (sum(diff < 0) + 1) / (length(diff) + 2))

p_value_one_sided <- bootstrap_accuracy_diff %>% 
  with(mean(diff > 0))

p_dens_diff <- bootstrap_accuracy_diff %>% 
  ggplot() + 
  aes(x = diff) +
  geom_density(alpha = .4, bw = "sj", fill = "lightgray") +
  labs(
    x = "Difference in Median Absolute Error",
    y = "Density"
  ) + 
  scale_fill_manual(values = palette) +
  scale_color_manual(values = palette) +
  theme(
    axis.text.y = element_blank(), axis.ticks.y = element_blank(),
    legend.position = "top", legend.title = element_blank()
  ) + 
  geom_vline(xintercept = 0, linetype = "dashed") + 
  annotate(
    "text", x = -35, y = 0.015, color = palette[2], 
    label = sprintf(
      "Bootstrapped median difference\np.value = %0.3f", p_value
    )
  )

p_dens <- bootstrap_accuracy %>% 
  ggplot() +
  aes(x = boot_med, fill = treatment_explain_id, color = treatment_explain_id) +
  geom_density(alpha = 0.4, bw = "nrd") +
  labs(
    x = "Median Absolute Error",
    y = "Density (bootstrap)"
  ) +
  scale_fill_manual(values = palette) +
  scale_color_manual(values = palette) +
  theme(
    axis.text.y = element_blank(), axis.ticks.y = element_blank(),
    legend.position = "top",
    legend.title = element_blank()
  )

# Combine plots
(p_dens + p_dens_diff) + 
  plot_annotation(
    tag_levels = "A",
    caption = "Source: Full sample (N = 121, excluding one outlier)."
  )
```

To examine the impact of AI explanations on students' perceived credibility of AI predictions (RQ2), we analysed students' five-point ratings of the perceived accuracy of the AI estimated value of coins. We noticed that 53% of students who did not receive step-by-step reasoning considered the AI estimate as either “good” (39%) or “very good” (14%) compared to 43% of students exposed to the AI reasoning (19% and 24%, respectively). This gap of ten percentage points suggests that students viewed the AI as more accurate when the AI-generated guess was presented as a "black box." However, we didn't have enough observations to reach a statistically significant association (Fisher's test, p = .27), and even regression analysis, controlling for individual characteristics, showed no significant association (Figure \@ref(fig:perceived-accuracy-coins)).

```{r, include = FALSE, tab.cap = "Percentage of students who perceived the accuracy of the AI-generated estimate for the coin value in the jar as accurate"}
xtab <- xtabs(~ q10_guess_coins_ai_accuracy + treatment_explain_id, tutorgpt)
kable(prop.table(addmargins(xtab, 2), margin = 2) * 100, digits = 0)
```

```{r, include = FALSE, tab.cap = "% of students who perceived the accuracy of the Human-generated average guess for the coin value in the jar as _____"}
xtab <- xtabs(~ q11_guess_coins_human_accuracy + treatment_explain_id, tutorgpt)
kable(prop.table(xtab, margin = 2) * 100, digits = 0)
```

Conversely, exposure to the AI step-by-step reasoning seemed associated with students' perception of the accuracy of the "human" guess --- the average guess from 600 participants reported in the original study [@steiner2015turns].^[Notably, this guess was $596, exceeding the correct value ($379.54) by about the same amount as the AI guess underestimated it ($213).] While all groups received the same human and AI estimates, 64% of students in the control rated the human estimate as "poor" or "very poor" against 45% of students exposed to AI reasoning. Although this difference was insignificant (Fisher's exact test, p = 0.15),  regression analysis, controlling for fixed differences across schools and across individuals, such as their gender, the initial guess of the value of coins, and experience with ChatGPT, showed a significant effect (p < 0.05) of the assignment to AI reasoning on students' perceptions of the human guess, as shown in Figure \@ref(fig:perceived-accuracy-coins). This finding underscores the complex relationship between AI and learning. It suggests that students exposed to the AI's step-by-step reasoning may have identified minor errors, thus increasing their expectations of human accuracy, and those who viewed AI as a "black box" tended to underrate human estimates.

```{r perceived-accuracy-coins, fig.cap = "Comparison of students' perceived accuracy of AI and human estimates across treatments on a five-point scale. Coefficients from separate ordinal logistic regressions controlling for students' location,  gender, and prior experience with ChatGPT. Positive coefficients indicate increased perceived accuracy associated with students' exposure to AI reasoning.", fig.width = 7, fig.height=3.5, echo = FALSE, warning = FALSE, message = FALSE}

reverse_order <- function(x) {
  factor(x, rev(levels(x)), ordered = TRUE)
}

# Fit model
model <- reverse_order(q11_guess_coins_human_accuracy) ~ treatment_explain_id +
  location + has_used_before + gender

fit <- MASS::polr(model, tutorgpt)
fit_ai <- update(fit, reverse_order(q10_guess_coins_ai_accuracy) ~ .)

list("Human" = fit, "AI" = fit_ai) %>% 
  lapply(broom::tidy, conf.int = TRUE) %>% 
  dplyr::bind_rows(.id = "depvar") %>% 
  filter(term == "treatment_explain_idexplain_yes") %>%
  ggplot() + 
  aes(
    color = "95% CI",
    x = estimate, y = depvar,
    xmin = conf.low, xmax = conf.high
  ) +
  scale_y_discrete(
    labels = c(Human = "Perceived human accuracy", 
               AI = "Perceived AI accuracy")
  ) +
  scale_color_manual(values = "navy") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_linerange(linewidth = 3, alpha = 0.2) + 
  geom_point(size = 3) +
  guides(
    colour = guide_legend(position = "inside")
  ) + 
  theme(
    legend.position.inside = c(0.1, 0.9),
    legend.title = element_blank(),
  ) + 
  labs(
    x = "Regression coefficient (ordinal logistic)", 
    y = "Dep. var.",
  )
```

## A Comparison of Socratic vs Non-Socratic AI

### Student-AI Interactions

Figure \@ref(fig:messages) shows the students' message length and frequency in the Socratic AI and Non-Socratic AI groups, allowing us to compare the student-AI interactions across treatment groups. As shown in the figure, Socratic AI students exchanged a median of 20 messages, significantly more than the eight messages by non-Socratic AI students (Wilcoxon test, p < 0.01). In addition, the Socratic AI’s messages were significantly shorter, with a median of 42 words, compared to 123 for the non-Socratic tutor (Wilcoxon test, p < 0.01). Socratic students also used fewer words, with two peaks: one at ten and the other at one word. This evidence supports the hypothesis that the Socratic tutor encouraged more engagement and interaction, resembling a relatively more genuine student-tutor talk.

```{r messages, fig.cap = "Comparison of message frequency and word count per message across Socratic and non-Socratic treatments. Top panel (**A**) shows differences in the frequency of messages exchanged, while the bottom panels (**B** ) depict the word counts per message for the AI tutor (left) and the students (right).", fig.width = 7, fig.height = 7, echo = FALSE}
palette <- c("#d95f02", "#7570b3")

# Step 1. count messages per users for Panel A
messages <- chats %>%
  dplyr::filter(!is.na(created_at)) %>%
  dplyr::count(location, user_id, treatment_tutor_id, sort = TRUE)

# Show KS-test differences
# wilcox.test(n ~ treatment_tutor_id, data = messages)

# Compute medians
df_med_messages <- messages %>%
  summarise(n = median(n), .by = c(treatment_tutor_id)) %>% 
  mutate(
    label = case_when(
      treatment_tutor_id == "socratic" ~ 
        paste0("Socratic median exchanges = ", n),
      TRUE  ~ paste0("Non-Socratic median exchanges = ", n)
    )
  )

# Step 2: compute med word counts for panel B
df_med_words <- chats %>%
  summarise(words = median(words),
            .by = c(treatment_tutor_id, role_label)) %>%
  mutate(
    label = case_when(
      treatment_tutor_id == "socratic" ~ 
        paste0("Socratic med. words = ", words),
      TRUE  ~ paste0("Non-Socratic med. words = ", words)
    )
  )

# Step 3: Create Panel A - Message frequency plot
p_messages <- ggplot(messages) +
  aes(
    x = n,
    y = treatment_tutor_id,
    color = treatment_tutor_id,
    fill = treatment_tutor_id
  ) + 
  geom_boxplot(
    width = .15,
    fill = "whitesmoke",
    outlier.shape = NA
  ) +
  ggdist::stat_halfeye(
    adjust = .75,
    alpha = .25,
    width = 1,
    justification = -.2,
    point_colour = NA,
    point_interval = NULL
  )  +
  geom_point(
    size = .75,
    alpha = .5,
    position = position_jitter(width = .05, height = 0.05)
  ) +
  geom_text(
    data = df_med_messages,
    nudge_y = .25,
    aes(label = label),
    size = 3,
    fontface = "bold"
  ) +
  scale_y_discrete(
    labels = c(socratic = "Socratic", non_socratic = "Non-Socratic")
  ) + 
  scale_x_log10() +
  scale_fill_manual(values = palette) + 
  scale_color_manual(values = palette) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none",
    panel.spacing = unit(3, "lines"),
  )

# Stpe 4: Create Panel B
p_words <- chats %>% 
  ggplot() +
  aes(
    x = words,
    y = treatment_tutor_id,
    color = treatment_tutor_id,
    fill = treatment_tutor_id
  ) + 
  geom_boxplot(
    width = .15,
    fill = "whitesmoke",
    outlier.shape = NA
  ) +
  ggdist::stat_halfeye(
    alpha = .25,
    width = .5,
    .width = 0,
    justification = -.2,
    point_colour = NA,
    point_interval = NULL
  )  +
  geom_point(
    size = .15,
    alpha = .3, 
    position = position_jitter(width = .05, height = 0.05)
  ) +
  geom_text(
    data = df_med_words,
    nudge_y = .25,
    aes(label = label), 
    size = 3,
    fontface = "bold"
  ) + 
  facet_wrap(. ~ role_label) + 
  scale_color_manual(values = palette) + 
  scale_fill_manual(values = palette) + 
  scale_y_discrete(
    labels = c(socratic = "Socratic", non_socratic = "Non-Socratic")
  ) + 
  scale_x_log10() + 
  theme(
    legend.position = "none",
    panel.spacing = unit(3, "lines"),
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(),
  ) + 
  coord_cartesian(clip = FALSE)

# Panels A and B - combined
p_combined <-
  (p_messages + labs(
    title = NULL, 
    subtitle = NULL, 
    caption = NULL, 
    x = "Messages Exchanged with AI Tutor (per student)",
    y = NULL,
  )) /
  (p_words + labs(
    title = NULL, 
    subtitle = NULL, 
    caption = NULL, 
    y = NULL,
    x = "Word Counts (per message)"
  ))

p_combined + plot_annotation(tag_levels = "A") 
```

### Students' Confidence in Their Responses

To test whether Socratic AI increased students' confidence in their answers (RQ4), we examined students' self-reported confidence levels at the end of each task. As shown in Figure \@ref(fig:confidence), the observed differences in confidence between the treatment groups were minimal. Specifically, 25% of Socratic students reported feeling "very confident," and 33% reported feeling "confident" compared to 21% and 38%, respectively, among non-Socratic students. These differences were not statistically significant, even with regression analysis controlling for variation across students and tasks (Figure \@ref(fig:confidence-regression)). Thus, our analysis found no evidence of treatment effects on students' confidence in their answers.

```{r confidence, fig.width = 7, fig.height = 7, fig.cap = "This figure shows the impact of Socratic AI on students' confidence levels in their performance across three different tasks. Despite some differences between the Socratic and Non-Socratic groups, we found no significant association between the treatment assignment and students' declared confidence levels."}
create_impact_figure <- function(data, measure) {
  # Flip and bind data
  data_long <- data %>%
    bind_rows(mutate(data, name = "All")) %>%
    xtabs(
      formula = as.formula(paste("~", measure, "+ treatment_tutor_id + name")),
      data = .,
      exclude = "Other"
    )

  # Run Fisher's tests and get p-values
  test_results <- apply(data_long, 3, fisher.test)
  p_values <- sapply(test_results, function(x) x$p.value)

  # Convert xtabs results to dataframe and add percentage calculations
  plot_df <- as.data.frame(data_long) %>%
    dplyr::mutate(percent = 100 * (Freq + 1) / (sum(Freq) + 5),
         .by = c(treatment_tutor_id, name)) %>%
    dplyr::mutate(
      treatment_tutor_id = case_when(
        treatment_tutor_id == "socratic" ~ "Socratic AI",
        TRUE ~ "Non-Socratic AI"
      ),
      name = case_match(name,
        "All" ~ sprintf("All tasks<br>p = %2.3f", p_values["All"]),
        "task1_water" ~ sprintf("Task 1: Guess<br>p = %2.3f", p_values["task1_water"]),
        "task2_media" ~ sprintf("Task 2: Essay<br>p = %2.3f", p_values["task2_media"]),
        "task3_sound" ~ sprintf("Task 3: Physics<br>p = %2.3f", p_values["task3_sound"])
      )
    )

  # Set color palette
  palette <- c("#ca0020", "#f4a582", "#d1e5f0", "#92c5de", "#0571b0")

  # Create the plot
  this_plot <- ggplot(plot_df) +
    aes(
      x = percent,
      label = paste0(round(percent), "%"),
      y = get(measure),
      hjust = ifelse(percent < 5, -1, 0),
      fill = get(measure),
      position = name,
    ) +
    facet_grid(name ~ treatment_tutor_id, scales = "free") +
    scale_fill_manual(values = palette) +
    geom_col(width = .8, show.legend = FALSE, position = position_dodge(1), alpha = .75) +
    geom_text(position = position_stack(.5)) +
    theme(
      axis.line.x = element_blank(),
      axis.ticks = element_blank(),
      axis.text.x = element_blank(),
      strip.text.y = ggtext::element_markdown(face = "plain"),
      panel.spacing = unit(2, "lines")
    ) +
    labs(
      x = NULL,
      y = NULL,
      caption = "Source: Full sample of student-task combinations (n = 364)"
    )

  return(this_plot)
}

# Invert confidence levels 
tutorgpt_wide <- tutorgpt_wide %>% 
  mutate(value = factor(value, rev(levels(value))))

p1 <- create_impact_figure(data = tutorgpt_wide, measure = "value")
print(p1)

```

```{r confidence-regression, fig.width = 7, fig.height = 4, fig.cap = "This figure shows the coefficients from separate ordinal logistic regressions on students' confidence in their answer on a five-point scale, controlling for students' self-efficacy per task and task fixed effect. Results show no significant association between the treatment assignment and students' confidence levels.", message=FALSE, warning=FALSE}

fit_all <- ordinal::clm(
  value ~ treatment_tutor_id + name + conf_value,
  data = tutorgpt_wide
)

fit_task_1 <- ordinal::clm(
  value ~ treatment_tutor_id + conf_value + location, 
  data = tutorgpt_wide, subset = name == "task1_water"
)

fit_task_2 <- ordinal::clm(
  value ~ treatment_tutor_id + location + gender + has_used_before, 
  data = tutorgpt_wide, subset = name == "task2_media"
)

fit_task_3 <- ordinal::clm(
  value ~ treatment_tutor_id + conf_value + location, 
  data = tutorgpt_wide, subset = name == "task3_sound"
)

df_coef <- list(
  all = fit_all, task1 = fit_task_1, 
  task2=fit_task_2, task3=fit_task_3
) %>% 
  lapply(broom::tidy, conf.int = TRUE) %>% 
  bind_rows(.id = "name")

df_coef %>% 
  filter(grepl("treat", term)) %>%
  ggplot() +
  aes(
    color = "95% CI",
    x = estimate, y = name,
    xmin = conf.low, xmax = conf.high
  ) + 
  geom_coef() +
  scale_y_discrete(
    labels = c(
      task1 = "Task 1 (guessing)", 
      task2 = "Task 2 (writing)", 
      task3 = "Task 3 (physics)",
      all = "All tasks"
    )
  ) +
  theme(
    legend.position.inside = c(.1, .9), legend.title = element_blank(),
    plot.title.position = "plot",
  ) +
  labs(
    x = "Regression coefficient (ordinal logistic)", 
    y = "Tasks",
  )
```

Further explorative regression analysis to examine potential treatment interactions shows that Socratic AI students with prior ChatGPT experience reported significantly less confidence (p < 0.1) than Non-Socratic AI students. This explorative finding suggests that experienced users may perceive new or unconventional AI tutoring methods as less effective or even counterproductive, indicating that encouraging the use of such AI tutoring tools among experienced ChatGPT students can be challenging.

### Perceived Helpfulness of AI Tutor

To test differences in students' perceptions of the AI tutor's helpfulness (RQ5), we examined students' ratings of helpfulness of the AI interaction at the end of each task. These ratings were on a five-point scale, from "Not at all helpful" to "Very helpful."[^mistake]

[^mistake]: Due to a coding issue, the scale used in Brussels and Seville differed slightly in the labels: Seville's students saw "Extremely helpful" whereas Brussels students saw "Very helpful." However, this issue has a minimal impact on the overall results, as results focus on the negative labels ("not at all helpful" or "not helpful") and they remain the same even after controlling for location effects in a regression. Additionally, open discussions held with students after the experimental session confirmed that they found the Socratic AI less helpful. This feedback, combined with the observed differences in negative labels, reinforces that the coding discrepancy has a minimal impact on the overall findings.

In both treatment groups, many students found interacting with the AI tutor “very helpful” or “helpful”, with 56% of the non-Socratic and 44% of the Socratic students, as shown in Figure \@ref(fig:helpfulness). However, the Socratic treatment group showed a bimodal distribution, with a substantial fraction of students (21%) finding the interaction "not at all helpful." The association between perceived helpfulness and treatment assignment was statistically significant (Fisher's exact test, p = 0.025), providing evidence that Socratic AI was less helpful to certain students. This association, however, was stronger for Task 1 and Taks 3 than Task 2, suggesting an association between the perceived AI's helpfulness and the type of task. Ordinal logistic regression accounting for individual student and task characteristics, including self-efficacy per task, revealed a statistically significant negative difference that corresponds to a drop of approximately 13 percentage points in perceived helpfulness associated with the Socratic AI, as illustrated in Figure \@ref(fig:helpful-regression). See Section \@ref(sec:regress-confidence) for more details.

```{r helpfulness, fig.width = 7, fig.height = 7, fig.cap = "This figure shows the impact of Socratic AI on students' perceived helpfulness of the AI tutor across three different tasks."}
p2 <- create_impact_figure(data = tutorgpt_wide, measure = "helpful")
print(p2)
```

```{r helpful-regression, fig.width = 7, fig.height = 4, fig.cap = "The impact of Socratic AI on students' confidence levels in their performance across three different tasks. Despite some differences between the Socratic and Non-Socratic groups, we found no significant association between the treatment assignment and students' declared confidence levels", message=FALSE, warning=FALSE}

fit_all <- MASS::polr(helpful ~ treatment_tutor_id + name, tutorgpt_wide)
fit_task_1 <- MASS::polr(helpful ~ treatment_tutor_id, tutorgpt_wide, subset = name == "task1_water")
fit_task_2 <- MASS::polr(helpful ~ treatment_tutor_id, tutorgpt_wide, subset = name == "task2_media")
fit_task_3 <- MASS::polr(helpful ~ treatment_tutor_id, tutorgpt_wide, subset = name == "task3_sound")

df_coef <- list(all = fit_all, task1 = fit_task_1, task2=fit_task_2, task3=fit_task_3) %>% 
  lapply(broom::tidy, conf.int = TRUE) %>% bind_rows(.id = "name")

df_coef %>% 
  filter(grepl("treat", term)) %>%
  ggplot() +
  aes(
    color = "95% CI",
    x = estimate, y = name,
    xmin = conf.low, xmax = conf.high
  ) + 
  geom_coef() +
  scale_y_discrete(
    labels = c(
      task1 = "Task 1 (guessing)", 
      task2 = "Task 2 (writing)", 
      task3 = "Task 3 (sound propagation)",
      all = "All tasks"
    )
  ) +
  theme(
    legend.position.inside = c(.1, .9), legend.title = element_blank(),
    plot.title.position = "plot",
  ) +
  labs(
    x = "Regression coefficient (ordinal logistic)", 
    y = "Sample",
    title = "Effect of Socratic AI on Students' Perceived Helpfulness of the AI Tutor",
  )
```

### Learning Outcomes and Knowledge Retention

To evaluate the impact of learning (RQ3), we compared the correctness of responses to Task 3, which focused on sound propagation, and assessed knowledge retention using a follow-up question on the same topic without AI assistance. As shown in Figure \@ref(fig:learning), the use of AI significantly enhanced students' response accuracy. Before interacting with the AI tutor, only 32% of students correctly answered that sound travels faster in water than air due to water's higher density. After AI interaction, this percentage nearly doubled increasing to 68%. However, there was no significant difference in learning outcomes between the Socratic and non-Socratic AI approaches, as illustrated in the figure. Moreover, when presented with a follow-up question (asking in which medium sound travels fastest among gold, rubber, warm air, cold air, and water), only 18% of students responded correctly by selecting the denser material (i.e., gold), again with no significant difference across treatments. This result suggests two implications. Firstly, we found no evidence of Socratic AI improving learning. Secondly, our results seems to suggest a problem of limited retention of the insights obtained with AI assistance or difficulty in applying such learning to novel scenarios.[^limitation]

[^limitation]: While it is true that gold is significantly denser than water or air, it is possible that some students did not fully understand this fact or were unsure how to compare the densities of the materials. If that were the case, even if the AI effectively conveyed that sound travels faster in denser media, students lacking this knowledge may still have been unable to select the correct answer. However, this explanation seems unlikely, as the relative densities of air, water, and metals, like gold, are commonly taught and conceptually straightforward, suggesting that other factors may have contributed to students' outcomes.

```{r learning, fig.cap ="Effects on learning: **A** shows the % of correct responses about the physics of sound propagation, illustrating the positive effect of interacting with the AI tutor, while no differences are associated with the Socratic AI. **B** shows the results of the verification question asking students to identify the fastest material for sound propagation (speed as meter per second reported) without AI assistance. Only 18% responded accurately, indicating limited learning."}

# Verification question 
tutorgpt <- tutorgpt %>% 
  mutate(
    q29_sound_speed_verify = case_match(
      q29_sound_speed_verify,
      c("0", "very_helpful") ~ "Rubber",
      c("helpful", "1", "2") ~ "Warm/Cold Air",
      c("3", "neutral") ~ "Gold",
      c("4", "not_very_helpful") ~ "Water",
      .default = q29_sound_speed_verify
    )
  )

# Create table 1. 
tab <- tutorgpt %>% 
  bind_rows(., mutate(., treatment_tutor_id = "All")) %>% 
  xtabs(~ q25_sound_speed + treatment_tutor_id, .) %>%
  prop.table(2) * 100

colnames(tab) <- case_match(colnames(tab), 
                            "All" ~ "Full sample", 
                            "socratic" ~ "Socratic AI",
                            "non_socratic" ~ "Non-Socratic")

rownames(tab) <- case_match(rownames(tab), 
                            "same_speed" ~ "Same speed", 
                            "water_faster_higher_density" ~ "Water is faster because higher density (correct)",
                            "water_faster_lower_density" ~ "Water is faster because lower density",
                            "water_slower_higher_density" ~ "Water is slower because higher density",
                            "water_slower_lower_density" ~ "Water is slower because lower density",
                            .default = rownames(tab))

tab %>% 
  kbl(
    caption = "% of students in each treatment group by responses to the question 'Does sound travel faster in water than air?'",
    format = "pipe", digits = 1
  )

tutorgpt <- tutorgpt %>%
  mutate(
    q25_sound_speed_correct = ifelse(
      q25_sound_speed == "water_faster_higher_density", "correct", "incorrect"), 
    q26_sound_speed_with_ai_correct = ifelse(
      q26_sound_speed_with_ai ==  "water_faster_higher_density", "correct", "incorrect"),
    q26_update = ifelse(
      q25_sound_speed != q26_sound_speed_with_ai, "update_yes", "update_no")
  )

# tab <- tutorgpt %>%
#   bind_rows(., mutate(., treatment_tutor_id = "All")) %>% 
#   xtabs(~ q26_update + treatment_tutor_id, .) %>%
#   prop.table(margin = 2) * 100

# colnames(tab) <- case_match(colnames(tab), 
#                             "All" ~ "Full sample", 
#                             "socratic" ~ "Socratic AI",
#                             "non_socratic" ~ "Non-Socratic")

# rownames(tab) <- case_match(rownames(tab), 
#                             "update_yes" ~ "Update answer", 
#                             "update_no" ~ "No update answer",
#                             .default = rownames(tab))

# tab %>%
#   kbl(
#     format = "pipe", digits = 1,
#     caption = "% of students in each treatment group who changed their answers after interacting with the AI tutor"
#   )

# p_value <- with(tutorgpt, fisher.test(q26_update, treatment_explain_id))$p.value

# tab <- tutorgpt %>% 
#   bind_rows(., mutate(., treatment_tutor_id = "All")) %>% 
#   xtabs(~ q26_sound_speed_with_ai + treatment_tutor_id, .) %>%
#   prop.table(2) * 100

# colnames(tab) <- case_match(colnames(tab), 
#                             "All" ~ "Full sample", 
#                             "socratic" ~ "Socratic AI",
#                             "non_socratic" ~ "Non-Socratic")

# rownames(tab) <- case_match(rownames(tab), 
#                             "same_speed" ~ "Same speed", 
#                             "water_faster_higher_density" ~ "Water is faster because higher density (correct)",
#                             "water_faster_lower_density" ~ "Water is faster because lower density",
#                             "water_slower_higher_density" ~ "Water is slower because higher density",
#                             "water_slower_lower_density" ~ "Water is slower because lower density",
#                             .default = rownames(tab))

# tab %>% 
#   kbl(
#     caption = "% of students in each treatment group by responses to the question 'Does sound travel faster in water than air?' after interacting with the AI tutor",
#     format = "pipe", digits = 1
#   )

#' ## Verification question
#+ echo = FALSE
# tab <- tutorgpt %>% 
#   bind_rows(., mutate(., treatment_tutor_id = "All")) %>% 
#   xtabs(~ q29_sound_speed_verify + treatment_tutor_id, .) %>%
#   prop.table(2) * 100

# colnames(tab) <- case_match(colnames(tab), 
#                             "All" ~ "Full sample", 
#                             "socratic" ~ "Socratic AI",
#                             "non_socratic" ~ "Non-Socratic")

# rownames(tab) <- case_match(rownames(tab), 
#                             "same_speed" ~ "Same speed", 
#                             "water_faster_higher_density" ~ "Water is faster because higher density (correct)",
#                             "water_faster_lower_density" ~ "Water is faster because lower density",
#                             "water_slower_higher_density" ~ "Water is slower because higher density",
#                             "water_slower_lower_density" ~ "Water is slower because lower density",
#                             .default = rownames(tab))

# tab %>% 
#   kbl(
#     caption = "% of students in each treatment group by responses to the question 'Does sound travel faster in water than air?' after interacting with the AI tutor",
#     format = "pipe", digits = 1
#   )

strip_labels <- c(
  "(Intercept)" = "Non-Socratic AI", 
  "treatment_tutor_idsocratic" = "Diff. with Socratic AI"
)

# First panel 
panel_a <- tutorgpt %>% 
  pivot_longer(c(q25_sound_speed_correct, q26_sound_speed_with_ai_correct)) %>% 
  reframe(
    broom::tidy(lm(value == "correct" ~ treatment_tutor_id), conf.int = TRUE), .by = name
  ) %>%
  mutate(
    name = case_match(name, 
                      "q25_sound_speed_correct" ~ "Before AI",
                      "q26_sound_speed_with_ai_correct" ~ "After AI")
  ) %>%
  ggplot() +
  aes(x = estimate, y = name, xmin = conf.low, xmax = conf.high, color = "95% CI") +
  facet_grid(~ term, labeller = labeller(term = strip_labels)) + 
  geom_coef() + 
  geom_text_repel(aes(label = paste0(round(estimate * 100), "%"))) +
  theme(
    legend.position = "none",
    panel.spacing = unit(3, "lines"),
  ) +
  scale_x_continuous(labels = function(x) 100 * x) +
  labs(
    x = "Correct answers (%)",
    y = NULL,
    title = "Impact of AI on Answer Accuracy",
    subtitle = "Comparison of Correct Answers Before and After Interacting with AI Tutor"
  )

water <- data.frame(
  q29_sound_speed_verify = c("Gold", "Water", "Warm/Cold Air", "Rubber"),
  speed = c(3200, 1500, 350, 60)/1e3
)

panel_b_data <- tutorgpt %>%
  xtabs(~ q29_sound_speed_verify, .) %>% 
  as.data.frame() %>%
  left_join(water) %>% 
  mutate(
    percent = 100 * (Freq + 1) / (sum(Freq) + 4), 
    name = sprintf("%s<br>(%.1f km/s)", q29_sound_speed_verify, speed)
  )

panel_b <- panel_b_data %>% 
  ggplot() +
  aes(
    x = percent, y = reorder(q29_sound_speed_verify, speed),
    label = paste0(round(percent), "%"),
    hjust = ifelse(percent > 10, 1.2, -.2),
  ) +
  geom_col(fill = "orange") + 
  geom_text() +
  geom_text(aes(label = sprintf("%.1f m/s",speed)), x = 0, hjust = 0, color = "white") +
  theme(
    axis.text.y = element_markdown(lineheight = 1.3, hjust = 0)
  ) + 
  labs(
    x = "Students (%)", y = NULL
  )

(panel_a + labs(title = NULL, subtitle = NULL)) / (panel_b) + 
  plot_annotation(tag_levels = "A")
```