### 2025-06-03 HSSC_Revised_Submission/declaration.Rmd
edits

# Statements and Declarations {-}

## Competing Interests {-}

The authors declare no competing interests directly or indirectly related to the work submitted for publication.

## Declaration of generative AI and AI-assisted technologies in the writing process {-}

During the preparation of this work the author(s) used Grammarly, ChatGPT in order to improve the readability and language of the manuscript. After using this tool/service, the author(s) reviewed and edited the content as needed and take(s) full responsibility for the content of the published article.